---
title: "licorice"
output: 
  md_document
---

[![Build Status](https://travis-ci.org/Bart6114/licorice.svg?branch=master)](https://travis-ci.org/Bart6114/licorice)
[![Coverage Status](https://coveralls.io/repos/Bart6114/licorice/badge.svg?branch=master)](https://coveralls.io/r/Bart6114/licorice?branch=master)

```licorice``` is an R package that eases the plotting of _Likert-like_ data. It has been heavily inspired by the ```likert``` [package](https://github.com/jbryer/likert) from Bryer and Speerschneider.

```licorice``` makes use of the ```ggplot2``` plotting engine in such a way that interference by the ```licorice``` package in terms of theming is kept to a minimum (the graphs diplayed belowed are themed using the ```ggthemr``` package). When using the ```licorice``` function a ```ggplot2``` object is returned which can then be added upon to your liking. 

```{r setup, echo=FALSE, message=FALSE,warning=FALSE}
library(knitr)
opts_knit$set(upload.fun = imgur_upload, base.url = NULL)

library(ggthemr)
ggthemr("flat")
```

## Installation

For now, no CRAN version exists and you'll have to install from GitHub using ```devtools```.

```{r, eval=FALSE}
devtools::install_github("Bart6114/licorice")
```

## Preparing the data

The ```licorice``` function expects a given structure of data. The example ```pisatest``` dataset can be used as a reference.

A minimal requirement is the presence of the ```question```, ```response``` and ```count``` column. Additionally a ```group``` column can be added.

```{r, message=FALSE}
library(licorice)

head(pisatest)
```

For example the ```gapsample``` dataset is not structured as it should be. Some simple preparations make it suited for ```licorice```.

```{r}
head(gapsample)

library(dplyr)
gap_fixed<-
  gapsample %>%
  group_by(question, response) %>%
  summarise(count = n())

head(gap_fixed)

```


## Plotting the data
Three main plots are available. First a centered plot is shown; here the junction between two categories (which can be controlled by the ```middle_pos``` parameter) is centered. If the factor levels of the reponse variable are not set correctly, they can be specified using the ```answer_order``` parameter. If a ```middle_pos``` value of e.g. 2 is given, the results are centered at the junction between the second and third response type.

```{r, fig.height=10}
my_order<-
  c("Strongly disagree","Disagree", "Agree", "Strongly agree")


licorice(pisatest, answers_order = my_order, middle_pos = 2, type = "center", sort=T)
```

One can also fill the vertical space using a filled plot (also notice the ```sort``` argument).

```{r, fig.height=10}
licorice(pisatest, answers_order = my_order, type = "fill", sort=TRUE)
```

We can also have a look at the count data.


```{r, fig.height=10}
licorice(pisatest, answers_order = my_order, type = "count")
```

You can also show plots in combination with each other using existing functionality (the ```gridExtra``` library).

```{r, warning=FALSE, fig.height=10, fig.width=8, message=FALSE}
library(gridExtra)

grid.arrange(
  licorice(pisatest, my_order, middle_pos = 2.5, type = "center", sort=TRUE),
  licorice(pisatest, my_order, type = "count", sort=TRUE) +
    theme(axis.text.y=element_blank()) +
    scale_fill_discrete(""),
  ncol = 2,
  widths = c(3/4,1/4)
)
```

Groups (as the countries in the graph above) are shown automatically when a ```group``` column is available in the data set. For example, when using the generated ```gap_fixed``` data set (where not ```group``` column is present), no group is shown;

```{r, fig.width=8}
levels(gap_fixed$response)

licorice(gap_fixed, middle_pos = 4, sort=TRUE) +
  theme(legend.position="right")
```

